Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-01 main.tex
reorder sections
trying to unify approaches

Let $\Omega \subseteq \bbR^n$, $n \geq 1$, be a bounded connected open set. Given any $1 \leq p \leq \infty$, we let $L^{p}(\Omega)$ denote the Lebesgue space over $\Omega$ with integrability exponent $p$. Then $\bfL^{p}(\Omega) := L^{p}(\Omega)^{n}$ is the corresponding Lebesgue space of vector\cye{-valued} fields.

@@

Let $W^{1,p}(\Omega)$ denote the first-order Sobolev space over $\Omega$ with integrability exponent $p$, 
\begin{align} \label{equation:W1p}
    W^{1,p}(\Omega) := \left\{u \in L^{p}(\Omega) \cye{\suchthat} \nabla u \in \bfL^{p}(\Omega) \right\}.
\end{align}
For a vector-valued field $\bff$ that is a distributional gradient of an object from $W^{1,p}(\Omega)$, $\bff \in \nabla W^{1,p}(\Omega)$, we are \cye{interested in} {\em potentials}, i.e., scalar-valued functions $\Phi_{\cye{\grad}}( \bff ) \in W^{1,p}(\Omega)$ such that $\nabla \Phi_{\cye{\grad}}( \bff ) = \bff$. Then, 
%
\begin{align*}
    \Phi_{\cye{\grad}} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)
\end{align*}
is the {\em potential operator}. A particular example of our interest is
%
\begin{subequations}\label{math:pot_grad}\begin{align}\label{math:pot_grad_def}
    \Phi_{\cye{\grad}}( \bff ) := \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } \| u \|_{L^{p}(\Omega)}.
\end{align}
%
Actually, in the present gradient setting and since $\Omega$ is connected, already the gradient constraint determines $\Phi_{\cye{\grad}}( \bff )$ up to a constant\cye{, so that~\eqref{math:pot_grad_def} is a one-dimensional minimization problem}.

\cye{We want to see whether the $L^{p}(\Omega)$ norm of the potential $\Phi_{\cye{\grad}}( \bff )$ is bounded in terms of the $\bfL^{p}(\Omega)$ norm of the source field $\bff$, i.e., we want to evaluate the {\em operator norm}
\begin{align}\label{math:pot_grad_norm}
    C_{\grad,\Omega,p} := \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \Phi_{\cye{\grad}}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\cye{\bfL}^{p}(\Omega)} }.
\end{align}\end{subequations}}
This operator norm is indeed finite and it equivalently turns out to be the best constant in the {\em Poincar\'e--Friedrichs inequality}: for every vector field $\bff \in \nabla W^{1,p}(\Omega)$, there exists $u \in W^{1,p}(\Omega)$ such that $\nabla u = \bff$ and 
\begin{subequations}\label{math:PF_gradient}\begin{align} \label{math:PF_gradient_1}
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} 
    \| \bff \|_{\cye{\bfL}^{p}(\Omega)}.
\end{align}
Actually, \eqref{math:PF_gradient_1} is further equivalent to the more conventional writing
\begin{align} \label{math:PF_gradient_2}
    \min_{ c \in \bbR } \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p}
    \| \nabla u \|_{\cye{\bfL}^{p}(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    .
\end{align}\end{subequations}
We call $C_{\grad,\Omega,p}$ the \emph{Poincar\'e--Friedrichs constant} for the gradient operator with exponent $p$. \cye{A concise writing of all~\eqref{math:pot_grad}--\eqref{math:PF_gradient} is
\begin{align} \label{math:pot_PF_grad}
    \min\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } 
    \| u \|_{L^{p}(\Omega)}
    &\leq 
    C_{\grad,\Omega,p}
    \| \bff \|_{\cye{\bfL}^{p}(\Omega)}.
\end{align}
for all $\bff \in \nabla W^{1,p}(\Omega)$.}

@@

Consider the vector-valued spaces
\begin{subequations}\label{equation:W_curl_div_p} \begin{align}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in \bfL^{p}(\Omega) \cye{\suchthat} \curl \bfu \in \bfL^{p}(\Omega) \right\}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in \bfL^{p}(\Omega) \cye{\suchthat} \divergence \bfu \in L^{p}(\Omega) \right\}
    .
\end{align}\end{subequations}
The members of these spaces are those vector fields whose distributional curls and divergences are \cye{respectively} in \cye{the} Lebesgue spaces \cye{$\bfL^{p}$ or $L^{p}(\Omega)$}.  
This only requires that certain sums of distributional partial derivatives are integrable, 
and hence these spaces are not classical Sobolev spaces of vector\cye{-valued} fields.

\cye{Let a vector-valued field $\bff$ that is a distributional curl of an object from $\bfW^{p}(\curl,\Omega)$, $\bff \in \curl \bfW^{p}(\curl,\Omega)$, and scalar-valued field $f$ that is a distributional divergence of an object from $\bfW^{p}(\divergence,\Omega)$, $f \in \divergence \bfW^{p}(\divergence,\Omega)$, be given.} 
We are interested in constants for inequalities such as 
\begin{subequations}\label{math:pot_PF_curl_div} \begin{align}
    \min\limits_{ \substack{ \cye{\bfu} \in \bfW^{p}(\curl,\Omega) \\ \curl \cye{\bfu} = \bff } } 
    \| \cye{\bfu} \|_{L^{p}(\Omega)}
    &\leq 
    C_{\curl,\Omega,p}
    \| \bff \|_{L^{p}(\Omega)} \label{math:pot_PF_curl}
    ,
    \\
    \min\limits_{ \substack{ \cye{\bfu} \in \bfW^{p}(\divergence,\Omega) \\ \divergence \cye{\bfu} = f } } 
    \| \cye{\bfu} \|_{L^{p}(\Omega)}
    &\leq 
    C_{\divergence,\Omega,p}
    \| f \|_{L^{p}(\Omega)} \label{math:pot_PF_div}
    ,
\end{align}\end{subequations}
and how these constants can be estimated by the operator norms of {\em linear} potential operators 
\begin{align*}
    &\Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \\
    &\Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
The fundamental difference \cye{with respect to Section~\ref{section:intro_grad}} is that the curl and divergence have infinite-dimensional kernels. In Section~\ref{section:intro_grad}, the kernel of the gradient is the one-dimensional space of constant functions. It is thus trivially complemented for all $p$, with a canonical choice of projection. By contrast, in the Banach space case \cye{$p \neq 2$}, it is not immediately evident that the kernels of the curl and divergence operators are complemented, and a canonical projection only exists here in the Hilbert setting $p=2$. \cye{Actually,} the existence of a bounded linear potential operator is \cye{here} a non-trivial fact even in the Hilbert space case \cye{$p = 2$}.

@@

For vector-valued fields, new challenges arise \cye{because of} the infinite-dimensional kernels. We, \cye{however, overcome these challenges and succeed in generalizing} the sequential construction of the potential operator for the gradient to vector potentials\cye{, including in particular} the curl and divergence. The basic inductive strategy remains the same. We start by constructing, say, a curl potential, over a single simplex. Having already defined a potential operator over a \cye{set of simplices that we term} a subtriangulation, we then construct a potential over a simplex \cye{neighboring through some $(n-1)$-dimensional faces}. However, the potential \cye{on the new simplex} might \cye{here} not have the same tangential trace along the interface between the new simplex and the previous ones. Actually, the passage through the triangulation has to be such that the intermediate subtriangulations are manifolds with boundary and the interface is a boundary submanifold of the intermediate subtriangulation and the new simplex. 
Then, we can define a well-posed \cye{mapping from the already visited simplices.} Repeating this procedure exhausts the original triangulation and eventually provides the desired potential.
These stronger conditions on the triangulations are exactly those of a so-called \emph{shellable} triangulations. Shellability of simplicial complexes and, more generally, of polytopal complexes is a well-established notion in discrete geometry and combinatorics, see, e.g., Kozlov~\cite{kozlov2008combinatorial} or Ziegler~\cite{ziegler1995lectures}.
With respect to our main interest, local \cye{patches} (stars) in 2D and 3D triangulations are shellable.

@@

\todo[inline]{MV: Discussion reorganized.}

\cye{In this section, we wish to explain why we employ linear potential operators to estimate the Poincar\'e--Friedrichs constants and we also address the difference of the Poincar\'e--Friedrichs inequality~\eqref{math:PF_gradient_2} with the Poincar\'e inequality where the mean value of $u$ is fixed to zero.} \cye{We then review Poincar\'e--Friedrichs inequalities in the gradient case setting of Section~\ref{section:intro_grad}, with the emphasis on know results on analytical upper bounds for the constants.} We \cye{in particular} take the opportunity to discuss convex domains as a special case. This survey serves as a building block in constructing computable constants over triangulated domains \cye{in a combinatorial way below} but we believe it \cye{is also} of independent interest.

@@

In the Banach setting $p \neq 2$, finding a norm-minimizing potential \cye{such as $\Phi_{\cye{\grad}}( \bff )$ of~\eqref{math:pot_grad_def} for a given gradient $\bff$} is a nonlinear operation. Its existence \cye{may actually} not be guaranteed unless $1 < p < \infty$. 
\cye{However}, any \textit{linear} potential operator $\overline\Phi_{\cye{\grad}} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)$
satisfying $\nabla \overline\Phi_{\cye{\grad}}( \bff ) = \bff$ for any $\bff \in \nabla W^{1,p}(\Omega)$ \cye{has the} operator norm that \cye{gives} the \cye{upper} bound 
\begin{align}\label{math:PF_P}
    \mwl{C_{\grad,\Omega,p} = } 
    \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \Phi_{\cye{\grad}}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\cye{\bfL}^{p}(\Omega)} }
    \leq 
    \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \overline\Phi_{\cye{\grad}}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\cye{\bfL}^{p}(\Omega)} } \cye{=: \overline C_{\grad,\Omega,p}}
    .
\end{align}
\cye{This explains our interest in the construction of} {\em linear potential operators}.

@@

\cye{In the setting of Section~\ref{section:intro_grad}, when $\Omega$ is of finite measure,} a natural \cye{(and unique, since the gradient potential constraint $\nabla \overline \Phi_{\cye{\grad}}( \bff ) = \bff$ determines $\overline \Phi_{\cye{\grad}}( \bff )$ up to a constant)} choice \cye{for a linear potential operator} is the average-free potential operator $\overline\Phi_{\cye{\grad}} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)$ that is prescribed by
\begin{align*}
    \overline\Phi_{\cye{\grad}}( \nabla u ) \cye{:}= u - u_{\Omega},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    ,
\end{align*}
where $u_\Omega$ denotes the average of $u$ \cye{over $\Omega$}, that is,
\begin{align*}
    u_{\Omega} := \vol(\Omega)^{-1} \int_{\Omega} u(x) \;dx.
\end{align*}
\cye{This potential operator can also implicitly be defined by}
\begin{align*}
    \overline \Phi_{\grad}( \bff ) 
    := 
    \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \int_\Omega u = 0 } } \| \nabla u - \bff \|_{\cye{\bfL}^{p}(\Omega)},
    \quad 
    \forall 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*} 
Its operator norm is the \cye{best} constant $\overline C_{\grad,\Omega,p} > 0$ in the {\em Poincar\'e inequality} (or also Poincar\'e--Wirtinger inequality)
\begin{align}\label{math:P_gradient}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    \overline C_{\grad,\Omega,p} \| \nabla u \|_{\cye{\bfL}^{p}(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    . 
\end{align}
\cye{Thus, rephrasing~\eqref{math:PF_P} in terms of~\eqref{math:PF_gradient_2} and~\eqref{math:P_gradient},} 
%
\begin{align*}
    \cye{C_{\grad,\Omega,p} \leq \overline C_{\grad,\Omega,p}.}
\end{align*}

Towards a converse inequality, 
let us first observe that the average of any $u \in W^{1,p}(\Omega)$ with $p < \infty$ satisfies the bound 
\begin{align} \label{math:aver_norm}
    \| u_\Omega \|_{L^{p}(\Omega)}^{p}
    = 
    \int_{\Omega} \left( \vol(\Omega)^{-1} \int_{\Omega} |u(x)| \;dx \right)^{p}
    \leq 
    \int_{\Omega} \vol(\Omega)^{-1} \int_{\Omega} |u(x)|^{p} \;dx
    = 
    \| u \|_{L^{p}(\Omega)}^{p}
    .
\end{align}
Here, we have used H\"older's or Jensen's inequality. 
In the case $p = \infty$, any $u \in L^{\infty}(\Omega)$ satisfies $\| \cye{u}_\Omega \|_{L^{\infty}(\Omega)} \leq \| u \|_{L^{\infty}(\Omega)}$. 
We conclude that taking the average is a projection within Lebesgue spaces with unit norm. 
The triangle inequality now shows that 
\begin{align*}
    \| u - u_\Omega \|_{L^{p}(\Omega)} 
    \leq
    2
    \| u \|_{L^{p}(\Omega)},
    \quad 
    \forall
    u \in L^{p}(\Omega)
    .
\end{align*}
\cye{Thus}, the Poincar\'e--Friedrichs inequality implies the Poincar\'e inequality with 
\begin{align*}
    \cye{C_{\grad,\Omega,p}} \leq 2 \cye{\overline C_{\grad,\Omega,p}}.
\end{align*}
In the special case $p=2$, since the average \cye{becomes the} orthogonal projection, and this improves to $\| u - u_\Omega \|_{L^{2}(\Omega)} \leq \| u \|_{L^{2}(\Omega)}$ for any $u \in L^{2}(\Omega)$. Hence,
\begin{align*}
    \cye{\overline C_{\grad,\Omega,2}} = \cye{C_{\grad,\Omega,2}}. 
\end{align*}

\cye{From~\eqref{math:aver_norm},} Stern's generalized projection estimate~\cite[Theorem~4.1, Remark~5.1]{stern2015banach} implies improved estimate for all $L^{p}$ spaces with $1 \leq p \leq \infty$ \cye{as}
\begin{align*}
    \| u - u_\Omega \|_{L^{p}(\Omega)}
    \leq 
    \min\left( 2, 2^{|2/p-1|} \right)
    \| u \|_{L^{p}(\Omega)}
    = 
    2^{|2/p-1|} 
    \| u \|_{L^{p}(\Omega)}
    ,
    \quad 
    \forall 
    u \in L^{p}(\Omega)
    ,
\end{align*}
\cye{where we have also used} $1 \leq 2^{|2/p-1|} \leq 2$ for $1 \leq p \leq \infty$.
The upper bound is only attained in the limit cases $p = 1$ and $p = \infty$, and the lower bound is only attained if $p = 2$, where we reproduce the improved estimate for the Hilbert space case.
We thus conclude 
\begin{align}\label{math:p_vs_pf}
    \cye{C_{\grad,\Omega,p}} \leq 2^{|2/p-1|} \cye{\overline C_{\grad,\Omega,p}}.
\end{align}
In summary, our notion of Poincar\'e--Friedrichs constant is equivalent to the notion of Poincar\'e constant in the literature, up to a numerical factor that only depends on $1 \leq p \leq \infty$ and that is at most $2$.

>>>>

Let $\Omega \subseteq \bbR^n$, $n \geq 1$, be a bounded connected open set. Given any $1 \leq p \leq \infty$, we let $L^{p}(\Omega)$ denote the Lebesgue space over $\Omega$ with integrability exponent $p$. Then $\bfL^{p}(\Omega) := L^{p}(\Omega)^{n}$ is the corresponding Lebesgue space of vector  fields.

@@

Let $W^{1,p}(\Omega)$ denote the first-order Sobolev space over $\Omega$ with integrability exponent $p$, 
\begin{align} \label{equation:W1p}
    W^{1,p}(\Omega) := \left\{u \in L^{p}(\Omega) \cye{\suchthat} \nabla u \in \bfL^{p}(\Omega) \right\}.
\end{align}
For a vector-valued field $\bff$ that is a distributional gradient of an object from $W^{1,p}(\Omega)$, $\bff \in \nabla W^{1,p}(\Omega)$, we are \cye{interested in} {\em potentials}, i.e., scalar-valued functions $\Phi_{\grad}( \bff ) \in W^{1,p}(\Omega)$ such that $\nabla \Phi_{\grad}( \bff ) = \bff$. Then, 
%
\begin{align*}
    \Phi_{\grad} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)
\end{align*}
is the {\em potential operator}. A particular example of our interest is
%
\begin{subequations}\label{math:pot_grad}\begin{align}\label{math:pot_grad_def}
    \Phi_{\grad}( \bff ) := \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } \| u \|_{L^{p}(\Omega)}.
\end{align}
%
Actually, in the present gradient setting and since $\Omega$ is connected, already the gradient constraint determines $\Phi_{\grad}( \bff )$ up to a constant\cye{, so that~\eqref{math:pot_grad_def} is a one-dimensional minimization problem}.

\cye{We want to see whether the $L^{p}(\Omega)$ norm of the potential $\Phi_{\grad}( \bff )$ is bounded in terms of the $\bfL^{p}(\Omega)$ norm of the source field $\bff$, i.e., we want to evaluate the {\em operator norm}
\begin{align}\label{math:pot_grad_norm}
    C_{\grad,\Omega,p} := \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \Phi_{\grad}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\fooL^{p}(\Omega)} }.
\end{align}\end{subequations}}
This operator norm is indeed finite and it equivalently turns out to be the best constant in the {\em Poincar\'e--Friedrichs inequality}: for every vector field $\bff \in \nabla W^{1,p}(\Omega)$, there exists $u \in W^{1,p}(\Omega)$ such that $\nabla u = \bff$ and 
\begin{subequations}\label{math:PF_gradient}\begin{align} \label{math:PF_gradient_1}
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} 
    \| \bff \|_{\fooL^{p}(\Omega)}.
\end{align}
Actually, \eqref{math:PF_gradient_1} is further equivalent to the more conventional writing
\begin{align} \label{math:PF_gradient_2}
    \min_{ c \in \bbR } \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p}
    \| \nabla u \|_{\fooL^{p}(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    .
\end{align}\end{subequations}
We call $C_{\grad,\Omega,p}$ the \emph{Poincar\'e--Friedrichs constant} for the gradient operator with exponent $p$. \cye{A concise writing of all~\eqref{math:pot_grad}--\eqref{math:PF_gradient} is
\begin{align} \label{math:pot_PF_grad}
    \min\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } 
    \| u \|_{L^{p}(\Omega)}
    &\leq 
    C_{\grad,\Omega,p}
    \| \bff \|_{\fooL^{p}(\Omega)}.
\end{align}
for all $\bff \in \nabla W^{1,p}(\Omega)$.}

@@

Consider the vector-valued spaces
\begin{subequations}\label{equation:W_curl_div_p} \begin{align}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in \bfL^{p}(\Omega) \cye{\suchthat} \curl \bfu \in \bfL^{p}(\Omega) \right\}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in \bfL^{p}(\Omega) \cye{\suchthat} \divergence \bfu \in L^{p}(\Omega) \right\}
    .
\end{align}\end{subequations}
The members of these spaces are those vector fields whose distributional curls and divergences are \cye{respectively} in \cye{the} Lebesgue spaces \cye{$\bfL^{p}$ or $L^{p}(\Omega)$}.  
This only requires that certain sums of distributional partial derivatives are integrable, 
and hence these spaces are not classical Sobolev spaces of vector  fields.

\cye{Let a vector field $\bff$ that is a distributional curl of an object from $\bfW^{p}(\curl,\Omega)$, $\bff \in \curl \bfW^{p}(\curl,\Omega)$, and scalar-valued field $f$ that is a distributional divergence of an object from $\bfW^{p}(\divergence,\Omega)$, $f \in \divergence \bfW^{p}(\divergence,\Omega)$, be given.} 
We are interested in constants for inequalities such as 
\begin{subequations}\label{math:pot_PF_curl_div} \begin{align}
    \min\limits_{ \substack{ \bfu \in \bfW^{p}(\curl,\Omega) \\ \curl \bfu = \bff } } 
    \| \bfu \|_{L^{p}(\Omega)}
    &\leq 
    C_{\curl,\Omega,p}
    \| \bff \|_{L^{p}(\Omega)} \label{math:pot_PF_curl}
    ,
    \\
    \min\limits_{ \substack{ \bfu \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfu = f } } 
    \| \bfu \|_{L^{p}(\Omega)}
    &\leq 
    C_{\divergence,\Omega,p}
    \| f \|_{L^{p}(\Omega)} \label{math:pot_PF_div}
    ,
\end{align}\end{subequations}
and how these constants can be estimated by the operator norms of {\em linear} potential operators 
\begin{align*}
    &\Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \\
    &\Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
The fundamental difference \cye{with respect to Section~\ref{section:intro_grad}} is that the curl and divergence have infinite-dimensional kernels. In Section~\ref{section:intro_grad}, the kernel of the gradient is the one-dimensional space of constant functions. It is thus trivially complemented for all $p$, with a canonical choice of projection. By contrast, in the Banach space case $p \neq 2$, it is not immediately evident that the kernels of the curl and divergence operators are complemented, and a canonical projection only exists here in the Hilbert setting $p=2$. \cye{Actually,} the existence of a bounded linear potential operator here is a non-trivial fact even in the Hilbert space case $p = 2$.

@@

For vector fields, new challenges arise \cye{because of} the infinite-dimensional kernels. We, \cye{however, overcome these challenges and succeed in generalizing} the sequential construction of the potential operator for the gradient to vector potentials\cye{, including in particular} the curl and divergence. The basic inductive strategy remains the same. We start by constructing, say, a curl potential, over a single simplex. Having already defined a potential operator over a \cye{set of simplices that we term} a subtriangulation, we then construct a potential over a simplex \cye{neighboring through some $(n-1)$-dimensional faces}. However, the potential \cye{on the new simplex} might \cye{here} not have the same tangential trace along the interface between the new simplex and the previous ones. Actually, the passage through the triangulation has to be such that the intermediate subtriangulations are manifolds with boundary and the interface is a boundary submanifold of the intermediate subtriangulation and the new simplex. 
Then, we can define a well-posed \cye{mapping from the already visited simplices.} Repeating this procedure exhausts the original triangulation and eventually provides the desired potential.
These stronger conditions on the triangulations are exactly those of a so-called \emph{shellable} triangulations. Shellability of simplicial complexes and, more generally, of polytopal complexes is a well-established notion in discrete geometry and combinatorics, see, e.g., Kozlov~\cite{kozlov2008combinatorial} or Ziegler~\cite{ziegler1995lectures}.
With respect to our main interest, local \cye{patches} (stars) in 2D and 3D triangulations are shellable.

@@

In this section, we wish to explain why we employ linear potential operators to estimate the Poincar\'e--Friedrichs constants and we also address the difference of the Poincar\'e--Friedrichs inequality~\eqref{math:PF_gradient_2} with the Poincar\'e inequality where the mean value of $u$ is fixed to zero. We then review Poincar\'e--Friedrichs inequalities in the gradient case setting of Section~\ref{section:intro_grad}, with the emphasis on know results on analytical upper bounds for the constants. In particular, we take the opportunity to discuss convex domains as a special case. This survey serves as a building block in constructing computable constants over triangulated domains in a combinatorial way below but we believe it is also of independent interest.

@@

In the Banach setting $p \neq 2$, finding a norm-minimizing potential such as $\Phi_{\grad}( \bff )$ of~\eqref{math:pot_grad_def} for a given gradient $\bff$ is a nonlinear operation. Its existence may actually not be guaranteed unless $1 < p < \infty$. 
However, any \textit{linear} potential operator $\overline\Phi_{\grad} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)$
satisfying $\nabla \overline\Phi_{\grad}( \bff ) = \bff$ for any $\bff \in \nabla W^{1,p}(\Omega)$ has the operator norm that \cye{gives} the upper bound 
\begin{align}\label{math:PF_P}
    \mwl{C_{\grad,\Omega,p} = } 
    \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \Phi_{\grad}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\fooL^{p}(\Omega)} }
    \leq 
    \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \overline\Phi_{\grad}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\fooL^{p}(\Omega)} } \cye{=: \overline C_{\grad,\Omega,p}}
    .
\end{align}
This explains our interest in the construction of {\em linear potential operators}.

@@

\cye{In the setting of Section~\ref{section:intro_grad}, when $\Omega$ is of finite measure,} a natural \cye{(and unique, since the gradient potential constraint $\nabla \overline \Phi_{\grad}( \bff ) = \bff$ determines $\overline \Phi_{\grad}( \bff )$ up to a constant)} choice \cye{for a linear potential operator} is the average-free potential operator $\overline\Phi_{\grad} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)$ that is prescribed by
\begin{align*}
    \overline\Phi_{\grad}( \nabla u ) \cye{:}= u - u_{\Omega},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    ,
\end{align*}
where $u_\Omega$ denotes the average of $u$ over the domain $\Omega$, that is,
\begin{align*}
    u_{\Omega} := \vol(\Omega)^{-1} \int_{\Omega} u(x) \;dx.
\end{align*}
This potential operator can also implicitly be defined by
\begin{align*}
    \overline \Phi_{\grad}( \bff ) 
    := 
    \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \int_\Omega u = 0 } } \| \nabla u - \bff \|_{\fooL^{p}(\Omega)},
    \quad 
    \forall 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*} 
Its operator norm is the best constant $\overline C_{\grad,\Omega,p} > 0$ in the {\em Poincar\'e inequality} (or also Poincar\'e--Wirtinger inequality)
\begin{align}\label{math:P_gradient}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    \overline C_{\grad,\Omega,p} \| \nabla u \|_{\fooL^{p}(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    . 
\end{align}
Thus, rephrasing~\eqref{math:PF_P} in terms of~\eqref{math:PF_gradient_2} and~\eqref{math:P_gradient}, 
\begin{align*}
    C_{\grad,\Omega,p} \leq \overline C_{\grad,\Omega,p}.
\end{align*}

Towards a converse inequality, 
let us first observe that the average of any $u \in W^{1,p}(\Omega)$ with $p < \infty$ satisfies the bound 
\begin{align} \label{math:aver_norm}
    \| u_\Omega \|_{L^{p}(\Omega)}^{p}
    = 
    \int_{\Omega} \left( \vol(\Omega)^{-1} \int_{\Omega} |u(x)| \;dx \right)^{p}
    \leq 
    \int_{\Omega} \vol(\Omega)^{-1} \int_{\Omega} |u(x)|^{p} \;dx
    = 
    \| u \|_{L^{p}(\Omega)}^{p}
    .
\end{align}
Here, we have used H\"older's or Jensen's inequality. 
In the case $p = \infty$, any $u \in L^{\infty}(\Omega)$ satisfies $\| u_\Omega \|_{L^{\infty}(\Omega)} \leq \| u \|_{L^{\infty}(\Omega)}$. 
We conclude that taking the average is a projection within Lebesgue spaces with unit norm. 
The triangle inequality now shows that 
\begin{align*}
    \| u - u_\Omega \|_{L^{p}(\Omega)} 
    \leq
    2
    \| u \|_{L^{p}(\Omega)},
    \quad 
    \forall
    u \in L^{p}(\Omega)
    .
\end{align*}
\begin{align*}
    \color{Emerald}
    \| u - u_\Omega \|_{L^{p}(\Omega)}^{p}
    \leq 
    2^{p-1} 
    \| u \|_{L^{p}(\Omega)}^{p}
    +
    2^{p-1} 
    \| u_\Omega \|_{L^{p}(\Omega)}^{p}
    \leq 
    2^{p} 
    \| u \|_{L^{p}(\Omega)}^{p}
    .
\end{align*}
\mwl{Via an application of the triangle inequality, }Thus, the Poincar\'e--Friedrichs inequality implies the Poincar\'e inequality with 
\begin{align*}
    C_{\grad,\Omega,p} \leq 2 \overline C_{\grad,\Omega,p}.
\end{align*}
In the special case $p=2$, taking the average is an orthogonal projection, 
and so this improves to $\| u - u_\Omega \|_{L^{2}(\Omega)} \leq \| u \|_{L^{2}(\Omega)}$ for any $u \in L^{2}(\Omega)$. 
Hence,
\begin{align*}
    \overline C_{\grad,\Omega,2} = C_{\grad,\Omega,2}. 
\end{align*}
\mwl{
This improvement follows from the orthogonality of constant functions to average-free functions in $L^2(\Omega)$,
and it also follows from the projection estimate (see, e.g.,~\cite{xu2003some}).
Stern's generalized projection estimate~\cite[Theorem~4.1,Remark~5.1]{stern2015banach} implies improved estimate for all $L^{p}$ spaces with $1 \leq p \leq \infty$:
since taking the average is a the projection onto the constants functions with unit norm, it now follows that 
}
From~\eqref{math:aver_norm}, Stern's generalized projection estimate~\cite[Theorem~4.1, Remark~5.1]{stern2015banach} implies improved estimate for all $L^{p}$ spaces with $1 \leq p \leq \infty$ as 
\begin{align*}
    \| u - u_\Omega \|_{L^{p}(\Omega)}
    \leq 
    \min\left( 2, 2^{|2/p-1|} \right)
    \| u \|_{L^{p}(\Omega)}
    = 
    2^{|2/p-1|} 
    \| u \|_{L^{p}(\Omega)}
    ,
    \quad 
    \forall 
    u \in L^{p}(\Omega)
    .
\end{align*}
Here, we have used $1 \leq 2^{|2/p-1|} \leq 2$ for $1 \leq p \leq \infty$.
The upper bound is only attained in the limit cases $p = 1$ and $p = \infty$, and the lower bound is only attained if $p = 2$, where we reproduce the improved estimate for the Hilbert space case.
We thus conclude 
\begin{align}\label{math:p_vs_pf}
    C_{\grad,\Omega,p} \leq 2^{|2/p-1|} \overline C_{\grad,\Omega,p}.
\end{align}
In summary, our notion of Poincar\'e--Friedrichs constant is equivalent to the notion of Poincar\'e constant in the literature, up to a numerical factor that only depends on $1 \leq p \leq \infty$ and that is at most $2$.

%%%%

\subsection{Old : Review of Poincar\'e--Friedrichs constants}\label{section:poincare}
